Algomusicology, ?, Profit

### 2016-04-02 paper.tex
Elaboration.

\begin{abstract}
    In this paper I propose a novel method for converting algomusicology into profit. While my particular example
    is tongue-in-cheek, my serious point is that algorithmic composition can help us deconstruct and disrupt
    traditional notions of musical authorship.
\end{abstract}

@@

Art critics, literary critics, Algomusicologists have a challenge in common with the Underpants Gnomes, we may have
collected a vast amount of underpants/art/users, but how are we to turn that into profit? Less glibly, is algomusicology
doomed to be a mere observer and explainer of musical practice, or might it become an active participant? Computer science is
able to influence industry, though that might take longer than we'd like \footnote{Alonzo Church developed the
lambda calculus in the 1930s. Java gained lambda expressions in 2015.}. Can algomusicology influence music?

The examples in this paper are all written in Clojure, dialect of Lisp. I also rely on Leipzig, a small music theory
library written in Clojure and Midje, for music theoretical functions, and Midje, a unit testing library, for assertions.

@@

Though storing such a string would require roughly a petabyte of memory, representing it indirectly as a computation
takes very little space at all. To put it in perspective, that's roughly the size all music recorded to date,
described in less than a half a tweet \footnote{Based on the iTunes library featuring approximately 50 million songs
of approximately 4 megabytes each, and assuming it represents approximately a fifth of all recorded music.}.

@@

(fact "The result-length is how long the string
       representation of the evaluated result is."
  (result-length (repeat 65 \G)) => 131)

(fact "The description-length is how long the
       string representation of the expression is."
  (description-length (repeat 65 \G)) => 13)
\end{verbatim}

The randomness of a description can be measured by comparing the length of the textual description and the length
of the resulting value.

@@

(fact "Kolmogorov randomness is the compression
       ratio between the description and the result."
  (randomness (repeat 65 \G)) => 13/131)
\end{verbatim}

@@

\begin{verbatim}
(defmacro definitionally [macro sym]
  (let [value (-> sym repl/source-fn read-string last)]
    `(~macro ~value)))

(fact "The definitionally macro lets us calculate on the
       definition of symbols."
  (definitionally description-length row-row) => 281
  (definitionally result-length row-row) => 2037
  (definitionally randomness row-row) => 281/2037)
\end{verbatim}

@@

(fact "We can construct all strings as a lazy
       sequence."
  (->> (library-of-babel) (take 5))
    => ["" " " "!" "\"" "#"]
  (nth (library-of-babel) 364645)
    => "GEB")
\end{verbatim}

@@

Borges appreciated this paradox. His anonymous narrator concludes that "On some shelf in some hexagon (men reasoned)
there must exist a book which is the formula and perfect compendium of all the rest"?????. Any of the many books that contain
our library-of-babel function match the narrator's description, though the books that also contain the rest of this paper
must surely provide more helpful context.

@@

The Champernowne word is a transcendental constant whose decimal expansion is the concatenation of the digits of the
natural numbers.

\begin{verbatim}
(defn decompose [n]
  (let [[remainder quotient] ((juxt mod quot) n 10)]
    (if (zero? quotient)
      [remainder]
      (conj (decompose quotient) remainder))))

(defn champernowne-word
  ([from]
   (->> (range)
        (map (partial + from))
        (mapcat decompose)))
  ([]
   (champernowne-word 0)))

(fact "The Champernowne word is defined by concatenating the
       natural numbers base 10."
  (->> (champernowne-word) (take 16))
    => [0 1 2 3 4 5 6 7 8 9 1 0 1 1 1 2])
\end{verbatim}

@@

>>>>

\begin{abstract}
In this paper I propose a novel method for converting algomusicology into profit. While my particular example
is tongue-in-cheek, it shows that algorithmic composition and analysis can help us deconstruct and disrupt
traditional notions of musical authorship.
\end{abstract}

@@

Art critics, literary critics, algomusicologists and Twitter have a challenge in common with the Underpants Gnomes - we may have
collected a considerable quantity of underpants/art/users, but how are we to turn that into profit? Less glibly, is algomusicology
doomed to be a mere observer and explainer of musical practice, or might it become an active participant? Computer science is
able to influence industry, though that might take longer than we'd like \footnote{Alonzo Church developed the
lambda calculus in the 1930s. Java gained lambda expressions in 2015.}. Can algomusicology influence music?\footnote{This paper
appears to be the first reference to algomusicology. I believe the term, which is a portmanteau of algorave and musicology,
has a self-evident definition.}

This paper will argue that it can. It will do so by citing examples of existing art works that are algorithmic in nature, and
also by using algomusicological concepts to create a work that challenges our present conception of musical authorship and
copyright.

The examples in this paper are all written in Clojure, dialect of Lisp. I also rely on Leipzig, a music theory library written
in Clojure and Midje, a unit testing library, for its fact notation.

@@

Though storing such a string would require roughly a petabyte of memory, representing it indirectly as a computation
takes very little space at all. To put it in perspective, that's roughly the size all music ever recorded to date,
described in less than a half a tweet\footnote{Based on the iTunes library featuring approximately 50 million songs
of approximately 4 megabytes each, and assuming it represents approximately a fifth of all recorded music. Tweets are
restricted to 140 characters.}.

@@

(fact
  (result-length (repeat 65 \G)) => 131
  (description-length (repeat 65 \G)) => 13)
\end{verbatim}

The randomness of a description can be measured by comparing the length of the textual description and the length
of the resulting value. If there is no way to describe a value more concisely than the value itself, we say that it is
random.

@@

(fact
  (randomness (repeat 65 \G)) => 13/131)
\end{verbatim}

@@

\begin{verbatim}
(defmacro definitional [macro sym]
  (let [value (-> sym repl/source-fn read-string last)]
    `(~macro ~value)))

(fact
  (definitional description-length row-row) => 281
  (definitional result-length row-row) => 2037
  (definitional randomness row-row) => 281/2037)
\end{verbatim}

@@

(fact
  (->> (library-of-babel) (take 3)) => ["" " " "!"]
  (nth (library-of-babel) 364645) => "GEB")
\end{verbatim}

@@

Borges appreciated this paradox. His anonymous narrator reports that on "some shelf in some hexagon (men reasoned)
there must exist a book which is the formula and perfect compendium of all the rest"?????. Any of the many books that contain
our library-of-babel function match the narrator's description, though the books that also contain the rest of this paper
must surely provide more helpful context.

@@

In Carl Sagan's novel 'Contact', he imagines that a message to humanity is encoded in the digits of pi. If we design an
appropriate encoding scheme, we could do a similar thing and treat the digits of pi as music. However, we would encounter
a similar problem to Borges' librarians - we suspect that every sequence of data we could possibly want is in there
somewhere, but we don't know where to look.

Fortunately, there are other transcendental numbers that are more predictable. The Champernowne word is a constant whose
decimal expansion is the concatenation of the digits of the natural numbers.

\begin{verbatim}
(defn decompose-into-digits [n]
  (let [[remainder quotient] ((juxt mod quot) n 10)]
    (if (zero? quotient)
      [remainder]
      (conj (decompose-into-digits quotient) remainder))))

(defn champernowne-word
  ([from]
   (->> (range)
        (map (partial + from))
        (mapcat decompose-into-digits)))
  ([]
   (champernowne-word 0)))

(fact
  (->> (champernowne-word) (take 16))
    => [0 1 2 3 4 5 6 7 8 9 1 0 1 1 1 2])
\end{verbatim}

@@

\section{Conclusion}
